Minimizing post-shock forecasting error using disparate information

### 2020-07-28 manuscript/manuscript.tex
changes to introductions, remarks discussions

We provide forecasting adjustment techniques with the goal of lowering overall forecast error when the time series under study has undergone a structural shock in real-time. Standard forecasting methods may not allow users to explore the consequences that structural shocks have on baseline forecasts \citep{baumeister2014real}. This is a general problem that has many real life applications. For example, one may be interested in forecasting the stock price of a company tomorrow after hearing terrible or great news about the company in after hours trading. Companies may be interested in forecasting the demand of their products to adjust production after they were involved in a brand crisis, but they only have recent sales data for which the company is operating well. All is not lost in this setting, one may be able to supplement the present forecast with past data borrowed from disparate time series which contain similar structural shocks. The core idea of our methodology is to sensibly aggregate similar past realized shock effects which arose from disparate time series, and then incorporate the aggregated shock effect estimator into the present forecast. Our method of combining disparate shock effects embraces ideas from conditional forecasting \citep{baumeister2014real, kilian2017structural}, time series pooling using cross-sectional panel data \citep{ramaswamy1993empirical, pesaran1999pooled, hoogstrate2000pooling, baltagi2008forecasting, koop2012forecasting, liu2020forecasting}, jforecasting with judgement and models \citep{svensson2005monetary, monti2008forecast}, synthetic control methodology \citep{abadie2010synthetic, agarwal2020two}, and expectation shocks \citep{croushore2006data, baumeister2014general, clements2019measuring}.

We study the post-shock forecasting problem in the context of additive shock effects in linear autoregressive models. In this post-shock forecasting setting, the researcher has a time series of interest which is known to have recently undergone a structural shock, and the post-shock response is not observed. The researcher must therefore move beyond the modeling paradigm that they were previously working under to accommodate this new shock \citep{monti2008forecast, svensson2005monetary}. One method for estimating the shock effect is to produce a conditional point forecast where a sequence of non-zero future structural shocks are conditioned upon and estimated \citep{baumeister2014real}. Such conditional point forecasts are appropriate when the shock sequence considered is within the range of historical experience \citep{kilian2017structural}. On the other hand, our methodology allows for the inclusion of outside data sources into this conditional forecasting context provided that the shock effects are all thought to arise from a similar data generating process.

@@

We estimate the shock effects that are present in the disparate time series for which post-shock responses are observed. We then aggregate these estimated shock effects and use this aggregated estimate as an estimator for the shock effect in the time series of interest. This estimator is then added to a forecast for the yet to be realized post-shock response corresponding to the time series of interest. shock effects in our post-shock forecasting framework is similar to ``expectation shocks'' which are studied in \cite{clements2019measuring}. The context in \cite{clements2019measuring} allowed for consistent estimation of expectation shocks under a vector autoregressive model, possibly involving an instrumental variable approach as in \cite{croushore2006data}. There is no consistency guarantee here. In our context, the yet to observed shock effect of interest is a realization from a random process that can only be partially explained by the realizations of the disparate time series. Therefore, the shock effect of interest can not be estimated consistently.

@@

\begin{remark}
  Our forecasting premise and estimation construction shares similarities with Bayesian viewpoints. From a Bayesian perspective, $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ can be interpreted as attempts to estimate the Bayes rule $\mathrm{E}\left(\alpha|\eta,\phi,\theta,\beta,\x,y,\delta,\gamma\right)$.
\end{remark}

@@

Moreover, the mean function for $\alpha_i$ under $\mc{M}_2$ can be extended to include more lagged predictors. Multiple shock effects can be nested within a time series; and time series in the donor pool can be dependent. As an example, we could consider a dependency structure for the September 2008 shock effects in our analysis of Conoco Phillips stock. But we note that consistency estimates from LOOCV with $k$ random draws may not work well if donor pool candidates are not mutually independent since the almost unbiased property hinges on the mutual independence among candidates in the donor pool.
%This generalization is possible since our risk-reduction propositions in Section \ref{properties} assumes the shock effect estimates are independent of $\alpha_1$, the shock effect for the time series of interest; and allows arbitrary dependence of shock effects in the donor pool. 
%Specific to this generalization, we edit the model of $\alpha_i$ to $\alpha_i = \mu_{\alpha} + \delta_i'\mathbf{x}_{3, T_i^* + 1} + \gamma_i'\mathbf{x}_{3, T_i^*} + \tilde{\varepsilon}_i$ for $i = 3, 4, 5$ and $\alpha_i = \mu_{\alpha} + \delta_i'\mathbf{x}_{4, T_i^* + 1} + \gamma_i'\mathbf{x}_{4, T_i^*} + \tilde{\varepsilon}_i$ for $i = 6$.
%This generalization was adopted in Section \ref{forecasting}. In this case, the variance expressions proposed in Section \ref{properties} will not work. However, the risk-reduction propositions in Section \ref{properties} will work  as long as those shock effects (not necessarily mutually independent) in the donor pool are independent of the one of interest. 
Although it is reflected in $\mc{M}_2$, we stress that our proposed methods allow $\alpha_i$ to follow arbitrary distributions provided that its first and second moments exist. The covariates in the model for $\alpha_i$ under $\mc{M}_2$ can be different from the covariates in the model of $y_{i,t}$. We also note that our post-shock framework can be extended to settings where the shock effect can be decomposed into separable estimable parts. An example of this is the additive shock effect estimators that we studied in our Conoco Phillips analysis.

@@

We have implicitly assumed that $\mathbf{W}^*$ is non-degenerate in the population.  Recall that in Section \ref{constructionofestimators} we noted that if there exists some $\mathbf{W}^*$ which satisfies \eqref{SCM} and $2p < n$, then there will be infinitely many solutions to $\mathbf{W}^*$. However, in applications, it is possible that $\mathbf{W}^*$ may take values on the boundary of $\mc{W}$, in which case bootstrapping may fail to estimate the distribution of $\hat{\alpha}_{\mrm{wadj}}$ \citep{andrews2000inconsistency}.  Moreover, when $2p < n$, $\mc{B}_u$ fails since the existence of infinitely many solutions is certain (if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM})), and will guarantee degeneracy of  $\mathbf{W}^*$. However, this issue will not occur under $\mc{B}_f$ since it takes $\mbf{W}^*$ as being fixed and the parameter space is $\Theta$ that does not involve the constrained $\mc{W}$. Nevertheless, it does not seriously compromise the inference according to our simulation results in Section 4 in the Supplementary Materials. Note that there are some philosophical distinctions between $\mc{B}_u$  and $\mc{B}_f$. $\mc{B}_u$ treats the donor pool as realizations from some infinite super-population of potential donors. In contrast, $\mc{B}_f$ treats the donor pool as being fixed  and known before the analysis is conducted, where the randomness comes from parameters and idiosyncratic error.

@@

 \textbf{Proof for statement (i):} Under $\mc{M}_1$, $\hat{\alpha}_{\rm adj}$ is an unbiased estimator of $\E{\alpha_1}$ because
  \begin{align*}
   E\left( \frac{1}{n}\sum_{i=2}^{n+1} \hat{\alpha}_i\right)
   = \frac{1}{n}\sum_{i=2}^{n+1}\E{\hat{\alpha}_i}
   &= \frac{1}{n}\sum_{i=2}^{n+1}\E{\E{\hat{\alpha}_i\mid \Theta}}\\
   &=  \frac{1}{n}\sum_{i=2}^{n+1}\E{\alpha_i}
   = \mu_{\alpha}=\E{\alpha_1},
  \end{align*}
  where we used the fact that OLS estimator is unbiased when the design matrix $\mbf{U}_i$ is of full rank for all $i = 2, \ldots, n+1$. Because $\alpha_1\indep \varepsilon_{i,t}$, $\E{\hat{\alpha}_{\rm adj}\alpha_1}=\E{\hat{\alpha}_{\rm adj}}\E{\alpha_1}=(\E{\hat{\alpha}_{\rm adj}})^2$. By Lemma \ref{risklemma}, 
    \begin{align*}
    R_{T_1^*+1,1}-R_{T_1^*+1,2}
    &=\E{\alpha_1^2}-\E{\hat{\alpha}_{\rm adj}-\alpha_1}^2\\
   & =\E{\alpha_1^2}-\E{\alpha_1^2}- \E{\hat{\alpha}_{\rm adj}^2}+2\E{\hat{\alpha}_{\rm adj}\alpha_1} \\
   &= \mu_{\alpha}^2 - \var{\hat{\alpha}_{\rm adj}} 
  \end{align*}
  Therefore, as long as we have $\var{\hat{\alpha}_{\rm adj}}<\mu_{\alpha}^2$, we will achieve the risk reduction.

>>>>

We provide forecasting adjustment techniques with the goal of lowering overall forecast error when the time series under study has undergone a structural shock in real-time. Standard forecasting methods may not allow users to explore the consequences that structural shocks have on baseline forecasts \citep{baumeister2014real}. This is a general problem that has many real life applications. For example, one may be interested in forecasting the stock price of a company tomorrow after hearing terrible or great news about the company  after hours trading. Companies may be interested in forecasting the demand of their products to adjust production after they were involved in a brand crisis, but they only have recent sales data for which the company is operating well. All is not lost in this setting, one may be able to supplement the present forecast with past data borrowed from disparate time series which contain similar structural shocks. The core idea of our methodology is to sensibly aggregate similar past realized shock effects which arose from disparate time series, and then incorporate the aggregated shock effect estimator into the present forecast. Our method of combining disparate shock effects embraces ideas from conditional forecasting \citep{baumeister2014real, kilian2017structural}, time series pooling using cross-sectional panel data \citep{ramaswamy1993empirical, pesaran1999pooled, hoogstrate2000pooling, baltagi2008forecasting, koop2012forecasting, liu2020forecasting}, forecasting with judgement and models \citep{svensson2005monetary, monti2008forecast}, synthetic control methodology \citep{abadie2010synthetic, agarwal2020two}, and expectation shocks \citep{croushore2006data, baumeister2014general, clements2019measuring}.

We study the post-shock forecasting problem in the context of additive shock effects in linear autoregressive models. In this post-shock forecasting setting, the researcher has a time series of interest which is known to have recently undergone a structural shock, and the post-shock response is not observed. The researcher must therefore move beyond the modeling paradigm under which he or she was previously working  to accommodate this new shock \citep{monti2008forecast, svensson2005monetary}. One method for estimating the shock effect is to produce a conditional point forecast where a sequence of non-zero future structural shocks are conditioned upon and estimated \citep{baumeister2014real}. Such conditional point forecasts are appropriate when the shock sequence considered is within the range of historical experience \citep{kilian2017structural}. On the other hand, our methodology allows for the inclusion of outside data sources and covariates into this conditional forecasting context provided that the shock effects are all thought to arise from a similar data generating process. For further differences of assumptions on shocks, our methodology allows for \emph{unprecedented} shocks and no observation of past shocks.

@@

We estimate the shock effects that are present in the disparate time series for which post-shock responses are observed. We then aggregate these estimated shock effects and use this aggregated estimate as an estimator for the shock effect in the time series of interest. This estimator is then added to a forecast for the yet to be realized post-shock response corresponding to the time series of interest. Shock effects in our post-shock forecasting framework is similar to ``expectation shocks'' which are studied in \cite{clements2019measuring}. The context in \cite{clements2019measuring} allowed for consistent estimation of expectation shocks under a vector autoregressive model, possibly involving an instrumental variable approach as in \cite{croushore2006data}. There is no consistency guarantee here. In our context, the yet to observed shock effect of interest is a realization from a random process that can only be partially explained by the realizations of the disparate time series. Therefore, the shock effect of interest can not be estimated consistently.

@@

\begin{remark}
  Our forecasting premise and estimation construction shares similarities with Bayesian viewpoints. From a Bayesian perspective, if we assign a prior $\pi$ to $\alpha_1$, $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ can be interpreted as the Bayes rules with respect to $\pi$ under different loss functions. If the sampling distribution of the data and $\pi$ are known, it is possible to compute the Bayes risks of $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ with respect to $\pi$, and thus enable comparisons among them. Additionally, from Lemma 20.3 \citep{marden}, $\hat{\alpha}_{\rm adj}$, $\hat{\alpha}_{\rm wadj}$, and $\hat{\alpha}_{\rm IVW}$ are admissible among the set of Bayes rules with respect to $\pi$ under corresponding loss functions.
\end{remark}

@@

Moreover, the mean function for $\alpha_i$ under $\mc{M}_2$ can be extended to include more lagged predictors. Multiple shock effects can be nested within a time series; and time series in the donor pool can be dependent. As an example, we could consider a dependency structure for the September 2008 shock effects in our analysis of Conoco Phillips stock. But we note that consistency estimates from LOOCV with $k$ random draws may not work well if donor pool candidates are not mutually independent since the almost unbiased property hinges on the mutual independence among candidates in the donor pool.
%This generalization is possible since our risk-reduction propositions in Section \ref{properties} assumes the shock effect estimates are independent of $\alpha_1$, the shock effect for the time series of interest; and allows arbitrary dependence of shock effects in the donor pool. 
%Specific to this generalization, we edit the model of $\alpha_i$ to $\alpha_i = \mu_{\alpha} + \delta_i'\mathbf{x}_{3, T_i^* + 1} + \gamma_i'\mathbf{x}_{3, T_i^*} + \tilde{\varepsilon}_i$ for $i = 3, 4, 5$ and $\alpha_i = \mu_{\alpha} + \delta_i'\mathbf{x}_{4, T_i^* + 1} + \gamma_i'\mathbf{x}_{4, T_i^*} + \tilde{\varepsilon}_i$ for $i = 6$.
%This generalization was adopted in Section \ref{forecasting}. In this case, the variance expressions proposed in Section \ref{properties} will not work. However, the risk-reduction propositions in Section \ref{properties} will work  as long as those shock effects (not necessarily mutually independent) in the donor pool are independent of the one of interest. 
Although it is reflected in $\mc{M}_2$, we stress that our proposed methods allow $\alpha_i$ to follow arbitrary distributions provided that its first and second moments exist. The covariates in the model for $\alpha_i$ under $\mc{M}_2$ can be different from the covariates in the model of $y_{i,t}$. Recall that the modeling equation of $\alpha_i$ includes present day and lagged covariates. It can be also generalized to the one with $q$ lags covariates. In that case, based same assumptions in Section \ref{constructionofestimators}, a sufficient condition for $\mbf{W}^*$ to be unique is $p(q+1) \geq n$. We also note that our post-shock framework can be extended to settings where the shock effect can be decomposed into separable estimable parts. An example of this is the additive shock effect estimators that we studied in our Conoco Phillips analysis.

@@

We have implicitly assumed that $\mathbf{W}^*$ is non-degenerate in the population.  Recall that in Section \ref{constructionofestimators} we noted that if there exists some $\mathbf{W}^*$ which satisfies \eqref{SCM} and $2p < n$, then there will be infinitely many solutions to $\mathbf{W}^*$. However, in applications, it is possible that $\mathbf{W}^*$ may take values on the boundary of $\mc{W}$, in which case bootstrapping may fail to estimate the distribution of $\hat{\alpha}_{\mrm{wadj}}$ \citep{andrews2000inconsistency}.  Moreover, when $2p < n$, $\mc{B}_u$ fails since the existence of infinitely many solutions is certain (if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM})), and will guarantee degeneracy of  $\mathbf{W}^*$. However, this issue will not occur under $\mc{B}_f$ since it takes $\mbf{W}^*$ as being fixed and the parameter space is $\Theta$ that does not involve the constrained $\mc{W}$. Nevertheless, it does not seriously compromise the inference according to our simulation results in Section 4 in the Supplementary Materials. In fact, one can avoid this issue by increasing lagged number 1 to $q$ in the modeling equation of $\alpha_i$ to some $q$ such that $p(q+1) \geq n$, mentioned previously.

Note that there are some philosophical distinctions between $\mc{B}_u$  and $\mc{B}_f$. $\mc{B}_u$ treats the donor pool as realizations from some infinite super-population of potential donors. In contrast, $\mc{B}_f$ treats the donor pool as being fixed  and known before the analysis is conducted, where the randomness comes from parameters and idiosyncratic error.

@@

 \textbf{Proof for statement (i):} Under $\mc{M}_1$, $\hat{\alpha}_{\rm adj}$ is an unbiased estimator of $\E{\alpha_1}$ because
  \begin{align*}
   \mrm{E}\left( \frac{1}{n}\sum_{i=2}^{n+1} \hat{\alpha}_i\right)
   = \frac{1}{n}\sum_{i=2}^{n+1}\E{\hat{\alpha}_i}
   &= \frac{1}{n}\sum_{i=2}^{n+1}\E{\E{\hat{\alpha}_i\mid \Theta}}\\
   &=  \frac{1}{n}\sum_{i=2}^{n+1}\E{\alpha_i}
   = \mu_{\alpha}=\E{\alpha_1},
  \end{align*}
  where we used the fact that OLS estimator is unbiased when the design matrix $\mbf{U}_i$ is of full rank for all $i = 2, \ldots, n+1$. Because $\alpha_1\indep \varepsilon_{i,t}$, $\E{\hat{\alpha}_{\rm adj}\alpha_1}=\E{\hat{\alpha}_{\rm adj}}\E{\alpha_1}=(\E{\hat{\alpha}_{\rm adj}})^2$. By Lemma \ref{risklemma}, 
    \begin{align*}
    R_{T_1^*+1,1}-R_{T_1^*+1,2}
    &=\E{\alpha_1^2}-\E{\hat{\alpha}_{\rm adj}-\alpha_1}^2\\
   & =\E{\alpha_1^2}-\E{\alpha_1^2}- \E{\hat{\alpha}_{\rm adj}^2}+2\E{\hat{\alpha}_{\rm adj}\alpha_1} \\
   &= \mu_{\alpha}^2 - \var{\hat{\alpha}_{\rm adj}} 
  \end{align*}
  Therefore, as long as we have $\var{\hat{\alpha}_{\rm adj}}<\mu_{\alpha}^2$, we will achieve the risk reduction.
